\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-26 OaklandSubmission/impl.tex
nit

\section{Implementation}
\label{sec:impl}
The  \tool compiler is written in Python and
compiles each of our benchmarks in under a second to C++ code that
makes calls to the ABY
library~\cite{aby}. We use the default ABY configuration for our
experiment \footnote{We set the security parameter to 128 and use OT
extension-based arithmetic multiplication triples generation.}. ABY
provides multi-threading support (for the offline phase of the
protocol), we leverage the support and use at most four threads in our
evaluation.

@@

We use an off-the-shelf solver
(SeaHorn~\cite{seahorn}) to bound check the array indices
($\Gamma\vDash e < n$
in ({\sc {T-Read}}) and ({\sc{T-Write}})),
Figure~\ref{fig:compile}). We take the \tool source program and
translate it as an input to the solver. The solver takes less
than a minute on our largest benchmark.

@@

\divya{
\begin{enumerate}
\item add all the expressions that our language supports such as mult, bit-wise and and anything else.
\item Say that we compile to source to intermediate and use inter to circuit of ABY.. We can decide if we want to say this.
\end{enumerate}
}

>>>>

\section{Implementation}
\label{sec:impl}
We discuss some  implementation details of \tool.
The  \tool compiler is written in Python and
compiles each of our benchmarks in under a second to C++ code that
makes calls to the ABY
library~\cite{aby}. We use the default ABY configuration for our
experiment \footnote{We set the security parameter to 128 and use OT
extension-based arithmetic multiplication triples generation.}. ABY
provides multi-threading support (for the offline phase of the
protocol), we leverage the support and use at most four threads in our
evaluation.

@@

We use an off-the-shelf solver
(SeaHorn~\cite{seahorn}) to bound check the array indices
($\Gamma\vDash e < n$
in ({\sc {T-Read}}) and ({\sc{T-Write}})),
Figure~\ref{fig:compile}). We take the \tool source program and
translate it as an input to the solver. The solver takes less
than a minute on our largest benchmark to verify that all the array
accesses are in-bounds.

@@

### 2017-10-26 OaklandSubmission/related.tex

\begin{tiret}
\item {\bf Function Description:} Tools at this level are a language (similar to \tool) where the programmer can specify the functionality to be securely implemented in a high level language and the compiler compiles the functionality to generate a secure computation protocol. Such languages do not require the programmer to have expertise in writing cryptographic protocols. All such existing languages in literature are built with only one backend: Fairplay's Secure Function Definition Language (SFDL)~\cite{fairplay,fairplaymp} compiles high-level C or Pascal like programs into boolean circuits that are then evaluated using garbled circuits~\cite{yao} for the two-party case and distributed garbled circuits~\cite{bmr} for the multi-party case~\cite{bmr}. The Secure Multiparty Computation Language (SMCL)~\cite{smcl} that compiles high-level Java like programs into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff}. The work of Wysteria~\cite{wysteria} provides a way of writing ``mixed-mode'' programs that combine normal or per-party, local computations, with ``secure'' (joint) computations repeatedly in a seamless manner. This work compiles programs into boolean circuits and uses Choi {\em et al.}~\cite{choi} implementation of the GMW~\cite{gmw} protocol as the backend. The work of Mitchell {\em et al.}~\cite{lambdaps} compiles their language into either a backend that works over arithmetic secret shares or over boolean secret shares using homomorphic encryption~\cite{gentry}. Finally, the work of ObliVM~\cite{oblivm} also shows how to incorporate programs that may read/write portions of the secret data, based on other secrets, without revealing access patterns (using techniques from oblivious RAM~\cite{oram1,oram2}). Their compute is through boolean circuits and the garbling technique. It is to be noted that none of the above works combine arithmetic and boolean computation together within the same program. 
\item {\bf Protocol Description:} Tools at this level of abstraction provide a set of commands that can be put together to describe efficient secure computation protocols. These tools are at a high level (such as making a call to execute secure computation of an AND gate or a MULT gate over shared inputs). The exact cryptographic protocols that are used as backends to implement these secure computation protocols need not be known to the programmer, but in order to generate efficient protocols for the entire functionality, the programmer must know how the functionality will be broken into components that make calls to various secure computation protocols for sub-functions. As an example, the work of ABY \cite{aby} falls into this category. While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa). While the work of Kerschbaum {\em et al.}~\cite{kos14} provides a heuristic to automatically select the protocol for various types of sub-computations, one would still need to manually write the program using the share conversion tools of ABY. Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute). As noted by ABY~\cite{aby}, the performance is much poorer than tools like ABY that combine boolean and arithmetic secure computation techniques due to very high conversion costs between homomorphic encryption and boolean shares as well as the cost of homomorphic encryption itself as the security parameter increases. This abstraction requires a fairly involved knowledge of cryptography from the programmer. Other examples include the VIFF framework~\cite{viff} for arithmetic computations, Sharemind~\cite{sharemind} (that provides information-theoretic secure 3-party boolean computation), and the Secure Multiparty Computation language~\cite{nsy04,securemc}. 
\item {\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer. While the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks. This abstraction requires the highest knowledge of cryptography from the programmer. Examples of such tools include L1~\cite{l1} and Qilin~\cite{qilin}.
\end{tiret}

@@

>>>>

\\\\
\noindent{\bf Function Description:} Tools at this level are a language (similar to \tool) where the programmer can specify the functionality to be securely implemented in a high level language and the compiler compiles the functionality to generate a secure computation protocol. Such languages do not require the programmer to have expertise in writing cryptographic protocols. All such existing languages in literature are built with only one backend: Fairplay's Secure Function Definition Language (SFDL)~\cite{fairplay,fairplaymp} compiles high-level C or Pascal like programs into boolean circuits that are then evaluated using garbled circuits~\cite{yao} for the two-party case and distributed garbled circuits~\cite{bmr} for the multi-party case~\cite{bmr}. The Secure Multiparty Computation Language (SMCL)~\cite{smcl} that compiles high-level Java like programs into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff}. The work of Wysteria~\cite{wysteria} provides a way of writing ``mixed-mode'' programs that combine normal or per-party, local computations, with ``secure'' (joint) computations repeatedly in a seamless manner. This work compiles programs into boolean circuits and uses Choi {\em et al.}~\cite{choi} implementation of the GMW~\cite{gmw} protocol as the backend. The work of Mitchell {\em et al.}~\cite{lambdaps} compiles their language into either a backend that works over arithmetic secret shares or over boolean secret shares using homomorphic encryption~\cite{gentry}. Finally, the work of ObliVM~\cite{oblivm} also shows how to incorporate programs that may read/write portions of the secret data, based on other secrets, without revealing access patterns (using techniques from oblivious RAM~\cite{oram1,oram2}). Their compute is through boolean circuits and the garbling technique. It is to be noted that none of the above works combine arithmetic and boolean computation together within the same program. 
\\\\
\noindent{\bf Protocol Description:} Tools at this level of abstraction provide a set of commands that can be put together to describe efficient secure computation protocols. These tools are at a high level (such as making a call to execute secure computation of an AND gate or a MULT gate over shared inputs). The exact cryptographic protocols that are used as backends to implement these secure computation protocols need not be known to the programmer, but in order to generate efficient protocols for the entire functionality, the programmer must know how the functionality will be broken into components that make calls to various secure computation protocols for sub-functions. As an example, the work of ABY \cite{aby} falls into this category. While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa). While the work of Kerschbaum {\em et al.}~\cite{kos14} provides a heuristic to automatically select the protocol for various types of sub-computations, one would still need to manually write the program using the share conversion tools of ABY. Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute). As noted by ABY~\cite{aby}, the performance is much poorer than tools like ABY that combine boolean and arithmetic secure computation techniques due to very high conversion costs between homomorphic encryption and boolean shares as well as the cost of homomorphic encryption itself as the security parameter increases. This abstraction requires a fairly involved knowledge of cryptography from the programmer. Other examples include the VIFF framework~\cite{viff} for arithmetic computations, Sharemind~\cite{sharemind} (that provides information-theoretic secure 3-party boolean computation), and the Secure Multiparty Computation language~\cite{nsy04,securemc}. 
\\\\
\noindent{\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer. While the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks. This abstraction requires the highest knowledge of cryptography from the programmer. Examples of such tools include L1~\cite{l1} and Qilin~\cite{qilin}.

@@

\divya{We need to cite this paper \cite{yao-pipe} for yao pipelining. This technique was also used in \cite{valeriaMatrix,oblivm}}

### 2017-10-26 OaklandSubmission/train.tex

\subsection{Matrix factorization}
\tool, as a language, is not tied to secure prediction and can express more general computations.
To demonstrate the expressiveness of \tool, we use it to implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies. The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm 1 in~\cite{valeriaMatrix}).
This algorithm is based on gradient descent and iteratively converges to a local minima.
We implement this algorithm in \tool.

To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
manipulates 64-bit variables. For $\mathcal{M}$ Nikolaenko et al. consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset. The time reported in~\cite{valeriaMatrix}
for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
The second stage involves sorting and gradient computations and is the heaviest stage.
The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.

\begin{table}
\begin{tabular}{c|c|c|c |c|c| c}
  Stage         &  LAN (s) & WAN (s) & Comm. (Mb)  & depth & \#Gates & LOC\\
\hline
1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
\hline
2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
\hline
3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
\hline
Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
\end{tabular}

@@

We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
However, this benchmark requires more lines of code than the previous benchmarks.
It is inconvinient to write Batcher's sort in an iterative language (about 450 lines of \tool).
The size of a recursive implementation would be much slower.
In the future, we would like to add support for functions and recursion.
However, the current programmer effort seems miniscule compared to the mammoth implementation effort
put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
crypto-backend to this benchmark.

>>>>

\subsection{Matrix factorization}
\tool, as a language, is not tied to secure prediction and can express more general computations.
To demonstrate the expressiveness of \tool, we use it to implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret. The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
This algorithm is based on gradient descent and iteratively converges to a local minima.
We implement this algorithm in \tool.

To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
manipulates 64-bit variables. For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al. consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset. The time reported in~\cite{valeriaMatrix}
for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
The second stage involves sorting and gradient computations and is the heaviest stage.
The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.

\begin{table}
\begin{tabular}{c|c|c|c |c|c| c}
  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
\hline
1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
\hline
2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
\hline
3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
\hline
Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
\end{tabular}

@@

We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
However, this benchmark requires more lines of code than the previous benchmarks
because of  Batcher's sort (450 lines of \tool code in each stage).
However, the current programmer effort seems miniscule compared to the mammoth implementation effort
put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
crypto-backend to this benchmark.
